---
title: "Autocatalytic set"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Toy code from ChatGTP: "script to simulate and experiment with autocatalytic reactions. You can use this as a starting point to explore autocatalytic sets further. It simulates autocatalytic reactions between molecules A, B, and C and prints the final concentrations of the molecules after a specified number of iterations. You can modify the code to experiment with different molecules and reactions as needed"


## Clear workspace and set directory

```{r setup}
rm(list=ls())
getwd()
setwd("~/Github/astrodorm/")
```

## Load packages and functions

```{r}
require("png")
require("dplyr")
require("grid")
```

## Toy code
```{r}

# Define a function to simulate autocatalysis
simulate_autocatalysis <- function(molecules, reactions, num_iterations) {
  for (iteration in 1:num_iterations) {
    # Select a random reaction
    reaction <- sample(reactions, 1)
    
    # Check if all reactants are available in sufficient quantities
    reactants_available <- all(sapply(reaction$reactants, function(reactant) reactant$concentration >= 1))
    
    if (reactants_available) {
      # Perform the reaction
      for (reactant in reaction$reactants) {
        reactant$concentration <- reactant$concentration - 1
      }
      for (product in reaction$products) {
        product$concentration <- product$concentration + 1
      }
    }
  }
}

# Define a function to print the concentrations
print_concentrations <- function(molecules) {
  for (molecule in molecules) {
    cat(paste(molecule$name, ": ", molecule$concentration, "\n"))
  }
}

# Define molecules
A <- list(name = "A", concentration = 100)
B <- list(name = "B", concentration = 100)
C <- list(name = "C", concentration = 0)

# Define reactions
reactions <- list(
  list(reactants = list(A, B), products = list(C)),
  list(reactants = list(B, C), products = list(A)),
  list(reactants = list(A, C), products = list(B))
)

# Simulate autocatalysis
num_iterations <- 1000
simulate_autocatalysis(list(A, B, C), reactions, num_iterations)

# Print the final concentrations
print_concentrations(list(A, B, C))
```